Well-Designed Environmental Markets Enable Large-Scale Marine Conservation

### 2019-09-08 docs/old_ms/Markets_Conservation_sept_04.tex
Moved docs around and finalized MS for submission to Nature

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
While it is commonly agreed that marine conservation should expand dramatically around the world, most countries have been unwilling to unilaterally undertake large-scale conservation in their own waters. When a country closes a significant fraction of its waters to fishing, it loses significant fishery revenue; while this may be more-than-made-up-for through spillover of protected fish, these benefits typically accrue to other countries or the high seas. In the context of expansive marine conservation in the Pacific, we show that thoughtful design of cross-country rights-based approaches to fisheries can actually incentivize, rather than relegate, large scale conservation. By combining a model of cross-country trading of rights
with vessel-level data before and after a large-scale conservation action is implemented, we show that two market design features are crucial.  First, only when trading of rights across countries is allowed, can a conservation-minded country capture the economic spillover benefits of their conservation actions. Second, the allocation of rights to countries must not be watered-down over time by a country's conservation efforts; for example, a country's annual allocation of rights should depend on the size of the fish stock in their waters, not on how hard it has fished. Overall, these results provide a 
template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

Recognizing a need to protect biodiversity and ecosystem services, various international bodies have committed to dramatically expand marine protected areas around the world \cite{oleary_2016}. Currently only about 3\% of the world's oceans are strongly protected \cite{sala_2018}, but these commitments call for 10\%-30\% of the ocean to be off-limits to fishing \cite{oleary_2016,dinerstein_2019}. To achieve these goals, huge swaths of sovereign nations' waters must be closed to fishing, yet few studies examine the incentives or economic mechanisms that would motivate a country to engage in such large-scale marine protection. Would any country rationally close 10\%, 30\%, or even 100\% of its waters to fishing if this means losing all fishing revenue from within the closed area? If all countries acted unilaterally, the answer is probably ``no''. Losing this important source of revenue would cripple many fishing-dependent economies and the spillover benefits from conservation would accrue to weakly regulated fisheries on the high seas. How would this answer change if fishing countries were able to capture the benefits of this spillover?  We find that international fishing effort markets, where nations trade the right to fish with each other, may provide a viable mechanism that incentivizes large-scale marine conservation. Here we show how the market can be designed (if new) or modified (if already existing) to incentivize large-scale marine conservation.

We are motivated by a real-world but relatively unknown institution called the Parties to the Nauru Agreement (PNA). Like an OPEC for tuna, the PNA is a coalition of nine Pacific island nations that collectively manages tuna fishing in its members' waters \cite{havice_2013,aqorau_2018}. These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Vessel-days are allocated across the nine nations, which then lease them to fishing vessels. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA approach to managing tuna fisheries and build on it to show how a market-based solution can be designed to incentivize conservation.

Not all market-based approaches to environmental management are created equal. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could lease the rights to catch the fish that spill-over, other countries would likely buy them, offsetting the costs of not fishing its own waters. But if the conserving country was not allowed to lease these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior.

@@

We first simulate a fishery where trading between countries is not allowed, and find that a spatial closure in Patch 1 will always result in a loss in revenues to Patch 1 (Fig. \ref{fig:PNA_model}A). Higher within-patch stock movement ($\theta$ = 0 implies no within-patch movement and $\theta = 1$ implies that fish are well-mixed within the fishing season) allows vessels to harvest the stock within the remaining open area, lowering the cost to Country 1. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock where fish can move in and out of the reserve, a spatial closure reduces the amount of biomass that is available for harvest in the conserving patch (\emph{i.e.} biomass outside the reserve), which reduces vessels' willingness to pay to fish in such waters (Fig. S1). When countries cannot trade, the costs of conservation are incurred by Patch 1, but the benefits are received by the eight other patches (revenues increase between 0\% and 6\%; Fig. S2) and the high seas.

How would trading between countries change these results? We simulate the same fishery, but now allow for vessel-days to be traded across patches. As before, a closure in Patch 1 lowers the value of vessel-days in that patch. But increased biomass in other patches causes prices in Patches 2 to 8 to increase. As a result, vessel-days from Patch 1 are traded to Patches 2 to 9, until prices are the same across all patches (Fig. S3). Under this market design, revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}B; note that the horizontal axis now only ranges from 0 to 0.8\% instead of from 0 to 100\% as in Fig. \ref{fig:PNA_model}A). With trading, the relative revenue drop will always be smaller than the relative effort drop, and the opposite is observed when there is no trading (Fig. S4). Overall, 88\% to 99\% of the costs of conservation can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

We have shown that trading significantly reduces the costs of conservation. However, a new question arises. How should rights be re-allocated every year once a country starts closing its waters to fishing? We test a range of allocation rules that weight effort ($\alpha$) and biomass ($1 - \alpha$) differently as the basis for rights allocation, and compare the resulting revenues to a fishery operating for 50 years without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in long-term losses to the conserving country of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, a biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in revenue losses as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biological features of a stock (\emph{e.g.} the biomass within a nation's waters), and not on fishing effort, the incentives for conservation are sustained through time.

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ, effectively displacing all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. We combine vessel-tracking data \cite{kroodsma_2018} and country-level license revenue data \cite{ffa_2017} to quantify the displacement of fishing effort and the likely costs of conservation. Of the 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 28 ``non-displaced vessels'' never fished in PIPA waters but fished within Kiribati's EEZ. The remaining 221 vessels were not continuously observed before and after but are included in our analyses, and we refer to these as ``other vessels''.

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
While it is commonly agreed that marine conservation should expand dramatically around the world, most countries have yet to undertake large-scale conservation in their own waters. When a country closes a large fraction of its waters to fishing, it may lose significant fishery revenue. While this may be more-than-made-up-for through spillover of protected fish, in the context of very large Marine Protected Areas these benefits typically accrue to other countries or the high seas. We show that thoughtful design of cross-country rights-based approaches to fisheries can actually incentivize, rather than relegate, large scale conservation. By combining a model of cross-country trading of rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that two market design features are crucial. First, only when trading of fishing rights across countries is allowed, can a conservation-minded country capture the economic spillover benefits of their conservation actions. Second, the allocation of rights to countries must not be watered-down over time by a country's conservation efforts; for example, a country's annual allocation of rights should depend on the size of the fish stock in their waters, not on how hard it has fished. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

Recognizing the need to protect marine biodiversity and ecosystem services, various international bodies have committed to dramatically expand marine protection around the world by making 30\% of the oceans off-limits to fishing \cite{oleary_2016,dinerstein_2019}. While most marine protected areas (MPAs) implemented to date have been small, it is widely-recognized that to meet these goals, very large MPAs must also be part of the strategy. But would any country rationally close 30\%, 80\% or even 100\% of its waters to fishing if this means losing all fishing revenue from within the closed area? At first glance, the answer is probably ``no''. Losing this important source of revenue would cripple many fishing-dependent economies. On the other hand, the closure may generate substantial ``spillover'' of fish and other benefits into adjacent waters which could, in principle, offset these losses. The problem with this argument is that for very large MPAs, these spillover benefits would accrue to other nations, with no obvious mechanism for the conserving country to recoup them. A viable solution may lie in international fishing effort markets, where nations trade the right to fish with each other. Here, we will show how fishing effort markets can be designed (if new) or modified (if already existing) to incentivize implementation of large-scale marine protected areas.

We are motivated by a real-world but relatively unknown institution called the Parties to the Nauru Agreement (PNA). Like an OPEC for tuna, the PNA is a coalition of nine Pacific island nations that collectively manages tuna purse seine fishing in its members' waters \cite{havice_2013,aqorau_2018}. These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how the market can be designed to incentivize conservation.

Not all market-based approaches to environmental management are created equal. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could lease the rights to catch the fish that spill-over, other countries would likely buy them, offsetting the costs of not fishing its own waters. But if the conserving country was not allowed to lease these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

@@

We first simulate a fishery where trading between countries is not allowed. This represents the status quo of any nation unilaterally engaging in large-scale conservation. Intuitively, we find that a spatial closure in Patch 1 will always result in a loss in revenues to Patch 1 (Fig. \ref{fig:PNA_model}A). Higher within-patch stock movement ($\theta$ = 0 implies no within-patch movement and $\theta = 1$ implies that fish are well-mixed within the fishing season) allows vessels to harvest the stock within the remaining open area, lowering the cost to Country 1. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock where fish can move in and out of the reserve, a spatial closure reduces the amount of biomass that is available for harvest in the conserving patch (\emph{i.e.} biomass outside the reserve), which reduces vessels' willingness to pay to fish in such waters (Fig. S1). When countries cannot trade, the costs of conservation are incurred by Patch 1, but the benefits are received by the eight other patches (revenues increase between 0\% and 6\%; Fig. S2) and the high seas. This first simulated fishery show the misalignment of incentives, where a conservation-minded country incurs large costs and provides public benefits to other nations for free.

How would trading between countries change these results? We simulate the same fishery, but now allow for vessel-days to be traded across patches. As before, a closure in Patch 1 lowers the value of vessel-days in that patch. But increased biomass in other patches causes prices in Patches 2 to 8 to increase. As a result, vessel-days from Patch 1 are traded to Patches 2 to 9, until prices are the same across all patches (Fig. S3). Under this market design, revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}B; note that the horizontal axis now only ranges from 0 to 0.8\% instead of from 0 to 100\% as in Fig. \ref{fig:PNA_model}A). With trading, the relative revenue drop will always be smaller than the relative effort drop, and the opposite is observed when there is no trading (Fig. S4). Overall, this shows that 88\% to 99\% of the costs of conservation can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

We have shown that trading significantly reduces the costs of conservation. However, a new question arises. How should rights be re-allocated every year once a country starts closing its waters to fishing? Customarily, the allocation of fishing effort rights is some formula that combines historical fishing effort and biomass in a nation's waters \cite{havice_2013}. We test a range of allocation rules that weight effort ($\alpha$) and biomass ($1 - \alpha$) differently as the basis for rights allocation. We simulate a fishery operating with closures for 50 years and compare the resulting revenues to a fishery without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in long-term losses to the conserving country of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, a biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in revenue losses as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biological features of a stock (\emph{e.g.} the biomass within a nation's waters), and not on fishing effort, the incentives for conservation are sustained through time.

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ, effectively displacing all fishing effort within its boundaries\cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat, PIPA may provide important recruitment and biomass benefits to adjacent waters \cite{hernndez_2019}. We combine vessel-tracking data \cite{kroodsma_2018} and country-level license revenue data \cite{ffa_2017} to quantify the displacement of fishing effort and the likely costs of conservation. Of the 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 28 ``non-displaced vessels'' never fished in PIPA waters but fished within Kiribati's EEZ. The remaining 221 vessels were not continuously observed before and after but are included in our analyses, and we refer to these as ``other vessels''.
